Dish: Distribution-oblivious Shell Scripting

### 2019-11-23 pldi/draft.tex

@@

The size of the script grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
The first number alone is interesting, as it captures \sys's output for the sequential script.
As there is no parallelism, it can be viewed as the initial cost corresponding to using \sys.
% this is the size of 
This change in size does not necessarily translate to observed runtime, as most of this size gain comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
FIFOs are simply pipes made, the in-kernel buffering and synchronization mechanisms are the same, leading to non-observable changes in runtime performance.

In terms of benchmarks, \ttt{grep} is a short script centered around an expensive \sta command.
% a DFA-based regular expression matching phase;
  while \ttt{grep} defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
The \ttt{sort} pipeline is equally short, but its central operation is a command in \pur.
The next two benchmarks, \ttt{wf} and \ttt{top-n}, are based on McIlroy's now classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
The \ttt{bi-gram} pipeline calculates n-grams of base two;
  it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.
\ttt{spell} is based on 
spell program~\cite{bentley1985spelling} 
The next pipeline is based on the original \ttt{spell} program developed by Johnson~\cite{bentley1985spelling}---another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in the dictionary.

Generally, the planning

    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/topn_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:topn}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/wf_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:wf}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/shortest_scripts_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:spell}
    \end{subfigure}%
    \caption{todo}
    \label{fig:microbenchmarks}
\end{figure*}

>>>>

    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/topn_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:topn}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/wf_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:wf}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/shortest_scripts_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:spell}
    \end{subfigure}%
    \caption{
      \textbf{Speedup achieved by \sys as a function of the level of parallelism (1--200).} 
      Three different configurations per benchmark:
      (i) distributed, focusing only on the distributed execution,
      (ii) +compile, adding \sys' pre-processing overheads, and
      (iii) +cat, adding a final merger step.
    }
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

The size of a script (col. 5) grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
The first number alone is interesting, as it captures \sys's output for the sequential script.
As there is no parallelism, it can be viewed as the initial cost corresponding to using \sys.
% this is the size of 
This change in size does not necessarily translate to observed runtime, as most of this size gain comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
FIFOs are simply pipes made, the in-kernel buffering and synchronization mechanisms are the same, leading to non-observable changes in runtime performance.

In terms of benchmarks, \ttt{grep} is a short script centered around an expensive \sta command.
% a DFA-based regular expression matching phase;
  while \ttt{grep} defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
The \ttt{sort} pipeline is equally short, but its central operation is a command in \pur.
The next two benchmarks, \ttt{wf} and \ttt{top-n}, are based on McIlroy's now classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
The \ttt{bi-gram} pipeline calculates n-grams of base two;
  it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.
The next pipeline is based on the original \ttt{spell} program developed by Johnson~\cite{bentley1985spelling}---another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in the dictionary.

% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents the speedup gained by \sys as a function of the level of parallelism (1--200).
% we experiment with a $200\times$ level of parallelism to see potential overheads
Each one of these plots reports on three different configurations:
  (i) distributed, which captures only the runtime of the distributed computation,
  (ii) compile + distributed, which adds the \sys' analysis and generation overhead, and
  (iii) cat + compile + distributed, which adds a final merger  gathering the results at the end of the pipeline.
The results show significant speedups, ranging between \todo{7--70$\times$}, depending on the distributability characteristics of individual pipelines.

Interestingly, the overhead from \sys's analysis and generation phases remains collectively below 1s even for the high-scalability cases.
We decided to push \sys's limits of these phases by synthesizing an artificial 1000-stage pipeline---a carefully-constructed, worst-case workload that is nowhere near the ones observed in practice (\S\ref{macro1}--\ref{macro3}).
We find the resulting times quite acceptable:
  \todo{8.23s} for parsing the script, 
  \todo{3.22s} for analysis, and 
  \todo{X.XXs} for rewriting the results of the analysis.
